Separating Use and Reuse to Improve Both

### 2017-05-18 FACS/paper.tex
new abstract

\begin{abstract}
In most object oriented languages subclassing/code-reuse implies
subtyping. This is considered by many a conceptual design error, but it
seems required for technical reasons, due to what we call the
\emph{leaking problem}. The leaking problem shows that the separating
code-reuse from subtyping is non-trivial and requires a significant
departure from the OO models in existing mainstream OO languages.

This paper synthesizes an OO language design for a nominally typed
language that allows code-reuse and subtyping to be decoupled using a
form of \emph{traits}. This design improves on previous work by
Damiani and Servetto by providing a simpler and more natural model for
OO programmers. By making our
type system distinguish between code use and code reuse we can avoid
many redundant abstract declarations that were required in previous
work by Servetto and Damiani, while supporting self construction,
binary methods and recursive types. Moreover, we provide a novel and elegant solution to uniformly
handle behavior and state within trait composition.

>>>>

\begin{abstract}
In most object oriented languages subclassing/inheritance implies
subtyping. This is considered by many a conceptual design error, but it
seems required for technical reasons, due to what we call the
\emph{this-leaking problem}, showing that separating
code-reuse from subtyping is non-trivial and requires a significant
departure from the OO models in existing mainstream OO languages.

We are aware of at least 3 independently designed research languages 
that address this limitation:Delta-Trait, Package Template, DeepFjig.

Leveraging on \emph{traits}, in this work we aim to synthesizes
the main ideas of those very different language designs.
Our result is a nominally typed minimalistic language, natural to OO programmers.
By making our type system distinguish between code-use and code-reuse
we can separe inheritance and subtyping, while avoiding 
redundant abstract declarations (required in Delta-Trait and DeepFjig), while supporting self construction,
binary methods and recursive types.
Moreover, we provide a novel and elegant solution to uniformly
handle behavior and state within trait composition.
